---
title: "Coding with R"
author: "Alex Takoudes"
date: "2023-01-12"
output: html_document
---
***
##### Markup Languages
* Use instructions to explain how a document should appear
* HTML
  + Most webpages use HTML. Fundamental
* LateX
  + Best for really complicated projects
  + Books/manuscripts
* Markdown
  + More friendly and easier to write

***
##### Coding with R
* R is helpful because it is *reproducible*
* Excel typically has more errors than R
* **Advantages**
  + Interactive use
  + Used for graphics and statistics
    + Creates better quality graphics
  + Very active community
  + Works on multiple platforms
* **Disadvantages**
  + Lazy evaluation
  + Packages are poorly documented
  + Functions are hard to learn
  + Unreliable packages
  + Problems with big data (multiple GBs)

***
##### Notes about coding
* Use the `<-` instead of `=` for assignment
* `Command + shift + c` hashtags out highlighted stuff
* If console shows a `+` instead of `>` it is waiting for something
  + Ex. did not finish parentheses
* Clicking `Run` only runs one line
  + You could highlight it all and click `Run`
  + `Source` button runs it all
  + `Source with echo` runs it all and shows you in console

##### Variables
* `camelCase` format starts with lower case and next word is uppercase
* `snake_case` format is a preferred method
* `avoid.periods` not used
* `.` reserved for a temporary variable
  + Not saved in the environment
  
##### Functions
* Blocks of code that perform a task
* You can use a short command over and over instead of writing code multiple times

***
##### R's Four Data Types

Dimensions | Homogenous | Heterogeneous
-----------|------------|--------------
**1-Dimension** | Atomic Vector | List
**2-Dimension** | Matric | Data Frame

***
##### Atomic Vectors
* One dimensional (a single row)
* Fundamental data structures in R programming
* Can be different classes
  + Character strings
  + Integers
  + Double (anything with a decimal)
  + Logical (Boolean)
  + Factor (categorizes, groups variable)

***
##### Adding names to vectors
```{r}
z <- runif(5) #randint
names(z) <- c("chow", "pug", "beagle", "greyhound", "akita")
print(z)
names(z) <- NULL
print(z)
```
***
##### NA Values (missing data)
```{r}

z <- c(3.2, 3.5, NA)
mean(z)

#check for NAs
anyNA(z)
is.na(z)
which(is.na(z))

```
***
##### Subsetting Vectors
* Filtering Vectors
* Vectors are indexed which is helpful

```{r}
z <- c(3.1, 9.2, 1.3, 0.4, 7.5)
z[1] #INDEX STARTS AT 1 NOT 0
z[c(4, 5)]  #return index 4 and 5
z[-c(2, 3)] #everything except 2 and 3
z[z == 7.5] #use double equals for exact match
z[z < 7.5]
names(z) <- letters[1:5] # subsetting with characters
z
subset(x = z, subset = z>1.5) #subset function
```

##### Functions

`sample`
```{r}
story_vec <- c("A", "Frog", "Jumped", "Here")
story_vec
sample(story_vec) #randomize
sample(story_vec, size = 3) #random sample size 3

story_vec <- c(story_vec[1], "Green", story_vec[2:4]) #adding to vector
story_vec

story_vec[2] <- "Blue" #replacing elements of vector
story_vec
```

`vec`
```{r}
vec <- vector() #creates the empty vector
vec <- vector(mode = "numeric", length = 5) #specific empty vector
vec
```

`rep`
```{r}
z <- rep(x=1:4, each=3)
z
```

`seq`
```{r}
z <- seq(from=2, to=4)
z
z <- seq(from=2, to=4, by=0.5)
z
```
