---
title: "Data Structures"
author: "Alex Takoudes"
date: "2023-01-17"
output: html_document
---

### Reviewing Assignment 3
```{r}
queue <- c("sheep", "fox", "owl", "ant")

#you can modify a specific index in the vector using []
queue[5] <- "serpent"
queue

#you can remove an index by using the "-"
queue <- queue[-5]
queue
```

***
## Data Structures
Dimensions | Homogeneous | Heterogeneous
-----------|------------|--------------
**1-Dimension** | Atomic Vector | List
**2-Dimension** | Matrix | Data Frame

### Vector Properties
* All atomic vectors are the same data type
* When you use `c()` to assemble different types, R **coerces** them
* c prioritizes different data types
  1. Logical (Boolean)
  2. Integer
  3. Double (floats)
  4. Character (strings)
```{r}
#Doubles (floats)
a <- c(2, 2.0)
typeof(a)
#
b <- c("purple", "green")
d <- c(a, b)
typeof(d)
```
#### Comparison operators yield a logical result
```{r}
a <- runif(10)
a > 0.5 #conditional statement

# How many elements in the vector are > 0.5
sum(a > 0.5)
```