---
title: "Analysis and Control Structures"
author: "Alex Takoudes"
date: "2023-01-30"
output: html_document
editor_options: 
  chunk_output_type: console
---

#### Setting up

```{r, collapse=T}
library(tidyverse)
library(ggthemes)
library(dplyr)
```

```{r}
# Data obtained from Veysey Powell and Babbitt 2016: An experimental test of buffer utility as a technique for managing pool-breeding amphibians https://datadryad.org/stash/dataset/doi:10.5061%2Fdryad.547rp

dryadData <- read.table("/Users/alextakoudes/BIOL1007A/Data/veysey-babbitt_data_amphibians.csv", header=TRUE, sep=",")

glimpse(dryadData)
```

## Analysis
#### Experimental designs
* Independent/explanatory variable vs. dependent/response variable
* Continuous (range of numbers: height, weight, temp) vs. discrete/categorical(categories: species, treatments, site)


x-axis/y-axis | continuous | categorical
-----------|------------|--------------
**continuous** | linear regression | logistic regression
**categorical** | T-test (2 groups) or Anova (2+) | chi-squared (count)

## Basic Linear Regression: Two continuous variables
#### Is there a relationship between the mean pool hydroperiod and the number of breeding frogs caught

```{r}

# model y (response) ~ x (explanatory)
regModel <- lm(count.total.adults ~ mean.hydro, data=dryadData)
summary(regModel)

#histogram looks normal
hist(regModel$residuals)

#bring out just the r squared value
summary(regModel)$"r.squared"

#scatterplot
regPlot <- ggplot(data=dryadData, aes(x=mean.hydro, y=count.total.adults)) +
  geom_point() +
  stat_smooth(method=lm, se=0.99) + #add linear regression line with error bars
  theme_few()

regPlot
```

## Basic Anova: Continuous vs. Categorical
#### Was there a statistically significant difference in the number of adults among wetlands?

```{r}
ANOmodel <- aov(count.total.adults~factor(wetland),data=dryadData) #factor(wetland) groups the numbered wetlands into categorical

summary(ANOmodel)

dryadData %>%
  group_by(wetland) %>%
  summarize(avgHydro = mean(count.total.adults, na.rm=T), n=n())
```

#### Boxplot
```{r}
dryadData$wetland <- factor(dryadData$wetland) #overwrite wetland data with factored wetland data
class(dryadData$wetland)

ANOplot <- ggplot(data=dryadData, aes(x=wetland, y=count.total.adults, fill=species)) +
  geom_boxplot() +
  scale_fill_grey()

ANOplot
# ggsave(file="SpeciesBoxplots.pdf", plot=ANOplot, device="pdf") save plot to pdf
```

## Logistic Regression: Categorical vs. Continuous
* gamma probabilities are best for continuous variables that are always positive and have a skewed distribution

```{r}
x_var <- sort(rgamma(n=200, shape=5, scale=5))
y_var <- sample(rep(c(1,0), each=100), prob=seq_len(200))
log_reg_data <- data.frame(x_var, y_var)

log_reg_model <- glm(y_var ~ x_var,      #generalized linear model
                     data=log_reg_data, 
                     family=binomial(link=logit)
                  )

summary(log_reg_model)

log_reg_plot <- ggplot(data=log_reg_data, aes(x=x_var, y=y_var)) +
  geom_point() +
  stat_smooth(method="glm", method.args=list(family=binomial))

log_reg_plot
```

## Contingency Tables (chi-squared): Categorical vs. Categorical
#### Are there differences in counts of males and females between species?


```{r}
count_data <- dryadData %>%
  group_by(species) %>%
  summarize(Males = sum(No.males, na.rm=T), 
            Females=sum(No.females, na.rm=T)
            ) %>%
  select(-species) %>% #get rid of names for the species
  as.matrix() #convert to matrix

count_data

row.names(count_data) =c("SS", "WF")
count_data

chisq.test(count_data)$residuals

# mosaic plot
mosaicplot(x=count_data, col=c("goldenrod", "gray"), shade=F)

# bar plot
count_data_long <- count_data %>%
  as_tibble() %>%  #need to make it a tibble in order to pivot longer
  mutate(Species=c(rownames(count_data))) %>%  #use the matrix to add species names
  pivot_longer(cols=Males:Females, names_to="Sex", values_to="Count")

count_data_long

plot <- ggplot(data=count_data_long, aes(x=Species, y=Count, fill=Sex)) +
  geom_bar(stat="identity", position="dodge") +
  scale_fill_manual(values=c("darkslategrey", "midnightblue"))

plot
```
***
# `if` Statements and `for` Loops
***
## `if` and `ifelse` statements
* Any final unspecified `else` captures the rest of the conditions
* `else` must appear on the same line as the expression
* Used for single values only
```{r, eval=F}
if(condition){expression1}

if(condition){expression1} else{expression2}
```

```{r}
z <- signif(runif(1), digits=2)

if (z>0.8){cat(z, "is a large number", "\n")}
```

