---
title: "Data Structures"
author: "Alex Takoudes"
date: "2023-01-17"
output:
  html_document: default
---

### Reviewing Assignment 3
```{r}
queue <- c("sheep", "fox", "owl", "ant")

#you can modify a specific index in the vector using []
queue[5] <- "serpent"
queue

#you can remove an index by using the "-"
queue <- queue[-5]
queue
```

***
## Data Structures
Dimensions | Homogeneous | Heterogeneous
-----------|------------|--------------
**1-Dimension** | Atomic Vector | List
**2-Dimension** | Matrix | Data Frame

### Vector Properties
* All atomic vectors are the same data type
* When you use `c()` to assemble different types, R **coerces** them
* c prioritizes different data types
  1. Logical (Boolean)
  2. Integer
  3. Double (floats)
  4. Character (strings)
```{r}
#Doubles (floats)
a <- c(2, 2.0)
typeof(a)
#
b <- c("purple", "green")
d <- c(a, b)
typeof(d)
```
#### Comparison operators yield a logical result
```{r}
a <- runif(10)
a > 0.5 #conditional statement

# How many elements in the vector are > 0.5
sum(a > 0.5)
```
#### Vectorization
* Add a constant to a vector
```{r}
z <- c(10, 20, 30)
z
z + 1 #performs element-wise addition of 1

y <- c(1, 2, 3)
z + y #results in an element by element operation on the vector
```
#### Recycling
* If vector lengths aren't equal
```{r}
z <- c(10, 20, 30)
x <- c(1, 2)
z + x
```
#### Simulating Data
* runif (Flat distribution)
* rnorm (Curved distribution)
```{r}
set.seed(123) #first set will always be the same numbers with seed
unif <- runif(n=100, min=5, max=10)
hist(unif)

rNN <- rnorm(100) #mean = 0, st.dev = 1
mean(rNN) #should be close to 0
hist(rNN)
```

***
## Matrix
* 2 dimensional (rows and columns)
* Homogeneous data type
* Is an atomic vector organized into rows and columns
```{r}
my_vec <- 1:12
m <- matrix(data = my_vec, nrow = 4) #builds it by columns by default
m
m <- matrix(data = my_vec, ncol = 3, byrow = T)
m
```

***
## Lists
* 1 dimensional (same as vectors)
* Can hold different data types

```{r}
myList <- list(1:10, matrix(1:8, nrow=4, byrow=T), letters[1:3], pi)
str(myList) #structure of list
```

#### Subsetting lists
* Using `[]` gives you a single item BUT not the elements
```{r}
myList <- list(1:10, matrix(1:8, nrow=4, byrow=T), letters[1:3], pi)
myList[4]
#myList[4] - 3 returns an error because it's still compartamentalized
typeof(myList[4])
#to grab object, use [[]]
myList[[4]] - 3
myList[[2]][4,1] #access the matrix, get the 4th row, 1st column

c(myList[[1]], myList[[2]]) #to obtain multiple elements within list
```

#### Name list items when they are created
```{r}
myList2 <- list(Tester = F, littleM = matrix(1:9, nrow=3))
myList2$Tester #calls the element by it's name

myList2$littleM[2,3] #access the matrix and pull the 2nd row, 3rd column
myList2$littleM[2,]  #access the matrix and pull the entire 2nd row
```

#### Unlist to string everything back to vector
```{r}
myList2 <- list(Tester = F, littleM = matrix(1:9, nrow=3))
unrolled <- unlist(myList2)
unrolled
```

### Data Analysis
* Practicing with plant data
```{r}
data(iris)
head(iris)
plot(Sepal.Length ~ Petal.Length, data = iris)
model <- lm(Sepal.Length ~ Petal.Length, data = iris)
results <- summary(model)
results

results$coefficients["Petal.Length", "Pr(>|t|)"]
unlist(results)$coefficients8 #pvalue in summary is always coefficients8
```
##### For linear regression we are interested in petal length to sepal length
* We care about the `Pr(>|t|)` value for `Petal.Length`

## Data Frames
* **Equal-lenghted** vectors, each which is a column
* `rbind()` and `cbind()` allow you to add rows and columns

```{r}
varA <- 1:12
varB <- rep(c("Con", "LowN", "HighN"), each=4)
varC <- runif(12)

dFrame <- data.frame(varA, varB, varC, stringsAsFactors=F)
dFrame

# Adding a row
new_data <- list(varA=13, varB="HighN", varC=0.668)
#rbind
dFrame <- rbind(dFrame, new_data)
dFrame

# Adding a column
new_var <- runif(13)
#cbind
dFrame <- cbind(dFrame, new_var)
head(dFrame)
```