---
title: "Data Frames and Functions"
author: "Alex Takoudes"
date: "2023-01-19"
output: html_document
---

## Finishing Matrices

#### Subsetting by Elements

```{r}
m <- matrix(data = 1:12, nrow=3)
m[1:2,] #just the first two rows
m[,2:4] #just 2-4 columns

```

#### Subset with logical (conditional statements)
* Select all columns for which totals are > 15
```{r}
colSums(m) #column sums
colSums(m) > 15 #logical true if greater than 15
m[,colSums(m) > 15] #indexes just the columns greater than 15

#rows sums
m[rowSums(m) == 22,]
```

#### Logical operators
* `==` `!=` `>` `<`

#### Subsetting to a vector changes the data type
```{r}
z <- m[1,] #subsets it to a vector
print(z)

z2 <- m[1, , drop=FALSE] #keeps it as a matrix
print(z2)

m2 <- matrix(data = runif(9), nrow=3)
m2[3,2]
```

#### Use the assignment operator to substitute values
```{r}
m2[m2<0.6]

data <- iris
tail(data)
data[3,2] #numbered indices
dataSub <- data[,c("Species", "Petal.Length")]
tail(dataSub)
```

#### Sort a data frame by values
```{r}
ordered_iris <- iris[order(iris$Petal.Length),] #order function goes before comma so it sorts the rows
```

***
## Functions in R
```{r}
functionName <- function(argx=1, argy=2){
  localVariable <- sample(1:10)
  return(localVariable)
}
```

#### Local Variables
* When you create variables in a function, they can only be accesed in that function
* We need a function to return a variable to be used in the environment
```{r}
myFunc <- function(a=3, b=4){
  z <- a + b
  return(z)
}
myFunc() #runs defaults
myFunc(b=3.4) #changes b value 
```

## Annotating Functions
```{r}
##################################
# Function: HardyWeinberg
# input: all allele frequency p (0,1)
# output: p and the frequencies of 3 genotypes AA AB BB
#-------------------------------------------------
HardyWeinberg <- function(p=runif(1)){
  if(p > 1.0 | p < 0.0){
    return("Function failure: p must be between 0 and 1")
  }
  q <- 1-p
  fAA <- p^2
  fAB <- 2*p*q
  fBB <- q^2
  vec_out <- signif(c(p=p, AA=fAA, AB=fAB, BB=fBB), digits=3)
  return(vec_out)
}
##################################

HardyWeinberg()
```
#### IF statements
* `|` represents 'or'
* Also needs the `{}` just like a function

#### Create a complex default value
```{r}
##################################################
# FUNCTION: fitLinear2       
# fits simple regression line
# inputs: list of predictor (x) and response (y)
# outputs: slope and p-value
#------------------------------------------------- 
fitLinear2 <- function(p=NULL){
  if(is.null(p)){ #no p value was inputted. Let's make one
    p <- list(x=runif(20), y=runif(20))
  }
  my_mod <- lm(p$x~p$y)
  my_out <- c(slope=summary(my_mod)$coefficients[2,1], 
              pValue=summary(my_mod)$coefficients[2,4]
              )
  plot(x=p$x, y=p$y)
  return(my_out)
}

fitLinear2()
my_pars <- list(x=1:10, y=runif(10))
fitLinear2(p=my_pars)
```