---
title: "Data Structures"
author: "Alex Takoudes"
date: "2023-01-17"
output: html_document
---

### Reviewing Assignment 3
```{r}
queue <- c("sheep", "fox", "owl", "ant")

#you can modify a specific index in the vector using []
queue[5] <- "serpent"
queue

#you can remove an index by using the "-"
queue <- queue[-5]
queue
```

***
## Data Structures
Dimensions | Homogeneous | Heterogeneous
-----------|------------|--------------
**1-Dimension** | Atomic Vector | List
**2-Dimension** | Matrix | Data Frame

### Vector Properties
* All atomic vectors are the same data type
* When you use `c()` to assemble different types, R **coerces** them
* c prioritizes different data types
  1. Logical (Boolean)
  2. Integer
  3. Double (floats)
  4. Character (strings)
```{r}
#Doubles (floats)
a <- c(2, 2.0)
typeof(a)
#
b <- c("purple", "green")
d <- c(a, b)
typeof(d)
```
#### Comparison operators yield a logical result
```{r}
a <- runif(10)
a > 0.5 #conditional statement

# How many elements in the vector are > 0.5
sum(a > 0.5)
```
#### Vectorization
* Add a constant to a vector
```{r}
z <- c(10, 20, 30)
z
z + 1 #performs element-wise addition of 1

y <- c(1, 2, 3)
z + y #results in an element by element operation on the vector
```
#### Recycling
* If vector lengths aren't equal
```{r}
z <- c(10, 20, 30)
x <- c(1, 2)
z + x
```
#### Simulating Data
* runif
* rnorm
```{r}
set.seed(123) #first set will always be the same numbers with seed
runif(n=5, min=5, max=10)
```