---
title: "dplyr Lecture"
author: "Alex Takoudes"
date: "2023-01-23"
output: html_document
---

## Tidyverse
* Collection of packages
* Shares philosophy, grammar (how code is structured), and data structures

## Operators
* Symbols that tell R to perform different operations b/w variables
* Assignment Operator: `<-`
* Arithmetic operators: `+` `-` `*` `/` `^` etc.
* Logical operators: `!` `&` `|` "not, and, or"
* Relational operators: `==` `!=` `>` `<` `<=` `>=`
* Miscellaneous operators:  
  + `%>%`: forward pipe
  + `%in%` : matching operator

## Dplyr
* New(er) package which provides a set of tools for manipulating datasets
* Specifically written to be fast
* Individual functions that correspond to common operations
* Conflicts with certain functions can be resolved
  + `dplyr::filter()`
```{r}
library(tidyverse) #library function to load in packages
```

#### Core verbs
1. `filter()`
2. `arrange()`
3. `select()`
4. `group_by()` `summarize()`
5. `mutate()`

#### Star wars dataset practice

```{r}
data(starwars)
class(starwars)
```

#### Tibble (`tbl`)
* Modern take on data frame
* Optimized dataframe

```{r}
glimpse(starwars) #clean way to look at a dataset
```

#### NAs
```{r}
anyNA(starwars)
starwars_clean <- starwars[complete.cases(starwars[,1:10]),]
glimpse(starwars_clean)
anyNA(starwars_clean)
```

#### `Filter()` Picks/subsets observations(*rows*) by their values

```{r}
filter(starwars_clean, gender=="masculine" & height<180 & height>100) #can filter by several conditions
```

#### `%in%` operator (matching)
* Similar to `==` but you can compare different length vectors
* Output depends on whatever the first vector is
```{r}
a <- LETTERS[1:10]
length(a)

b <- LETTERS[4:10]
length(b)

a %in% b #output depends on whatever the first vector is
b %in% a
```

#### Subsetting with `%in%`
```{r}
eyes <- filter(starwars, eye_color %in% c("blue", "brown"))
```

#### `arrange()` Re-orders rows
```{r}
arrange(starwars_clean, by=height) #defaults to lowest to highest
arrange(starwars_clean, by=desc(height)) #helper function desc() reverses order
arrange(starwars_clean, height, desc(mass)) #second variable "mass" used to break ties of height
sw <- arrange(starwars, by=height)
tail(sw) #puts all the NAs at the bottom
```

#### `select()` Chooses variables(columns) by their names
```{r}
select(starwars_clean, 1:10) #only certain number columns
select(starwars_clean, name:species) #up until a certain column
select(starwars_clean, -(height:sex)) #minus ignores certain columns
#everything() helper function is useful if you want to move some variables to the beginning
select(starwars_clean, name, gender, species, everything()) #rearrange columns
#contains() helper function picks the columns only with the condition
select(starwars_clean, contains("color"))
#ends_with()
#starts_with()
#num_range()
#select also renames columns
select(starwars_clean, haircolor=hair_color) #returns just the column with a new name
rename(starwars_clean, haircolor=hair_color) #returns all columns with renaming
```

***
#### `mutate()` creates new variables using functions of existing variables
```{r}
#height divided by mass column
mutate(starwars_clean, ratio=height/mass, .before=1) #puts it at the front
transmute(starwars_clean, mass, mass_lbs=mass*2.2, height) #only returns mutated columns
```

#### `summarize()`
```{r}
summarize(starwars_clean, mean_height=mean(height))
#n() helper function gives the number of observations
summarize(starwars_clean, mean_height=mean(height), total_number=n())
```

#### `group_by()`
```{r}
starwars_genders <- group_by(starwars, gender)
head(starwars_genders)
summarize(starwars_genders, mean_height=mean(height, na.rm=T), total_number=n()) #summarizes the groups
```