---
title: "dplyr Lecture"
author: "Alex Takoudes"
date: "2023-01-23"
output: html_document
---

## Tidyverse
* Collection of packages
* Shares philosophy, grammar (how code is structured), and data structures

## Operators
* Symbols that tell R to perform different operations b/w variables
* Assignment Operator: `<-`
* Arithmetic operators: `+` `-` `*` `/` `^` etc.
* Logical operators: `!` `&` `|` "not, and, or"
* Relational operators: `==` `!=` `>` `<` `<=` `>=`
* Miscellaneous operators:  
  + `%>%`: forward pipe
  + `%in%` : matching operator

## Dplyr
* New(er) package which provides a set of tools for manipulating datasets
* Specifically written to be fast
* Individual functions that correspond to common operations
* Conflicts with certain functions can be resolved
  + `dplyr::filter()`
```{r}
library(tidyverse) #library function to load in packages
```

#### Core verbs
1. `filter()`
2. `arrange()`
3. `select()`
4. `group_by()` `summarize()`
5. `mutate()`

#### Star wars dataset practice

```{r}
data(starwars)
class(starwars)
```

#### Tibble (`tbl`)
* Modern take on data frame
* Optimized dataframe

```{r}
glimpse(starwars) #clean way to look at a dataset
```

#### NAs
```{r}
anyNA(starwars)
starwars_clean <- starwars[complete.cases(starwars[,1:10]),]
glimpse(starwars_clean)
anyNA(starwars_clean)
```

#### `Filter()`
* Picks/subsets observations(*rows*) by their values

```{r}
filter(starwars_clean, gender=="masculine" & height<180) #can filter by two conditions
```