---
title: "`ggplot2`"
author: "Alex Takoudes"
date: "2023-01-24"
output: html_document
---

## Indtroduction to ggplot

#### Setup
```{r}
library(ggplot2)
library(ggthemes)
library(patchwork)
library(dplyr)
library(viridis)
```

#### Template for ggplot code
`p1 <- ggplot(data = <DATA>, mapping = aes(x = xVar, y = yVar)) + <GEOM FUNCTION>`

`print(p1`

* `aes` is aesthetics
* `<GEOM FUNCTION>`
  + `geom_boxplot`

## Qplot: Quick Plotting
* `I()` is important for making the colors actually work
* Quick method
* Uses vectors
```{r}
d <- mpg
glimpse(d)

#histogram
qplot(x=d$hwy, fill=I("green"), color=I("black"))

#scatter plot
qplot(x=d$displ, y=d$hwy, geom=c("smooth", "point"), method="lm") #geom adds the line

#box plot
qplot(x=d$fl, y=d$cty, geom="boxplot", fill=I("forestgreen"))

#bar plot
qplot(x=d$fl,geom="bar",fill=I("green"))
```

#### Creating Data
```{r}
x_trt <- c("Control","Low","High")
y_resp <- c(12,2.5,22.9)
qplot(x=x_trt, y=y_resp, geom="col", fill=I(c("khaki", "slategray", "goldenrod")))
```

## `ggplot` uses data frames instead of vectors
* [Themes:](https://ggplot2.tidyverse.org/reference/ggtheme.html)
  + base
  + bw
  + classic
  + linedraw
  + dark
  + minimal
```{r}
p1 <- ggplot(data=d, mapping=aes(x=displ, y=cty, color=cyl)) +  #good style to indent after the plus
  geom_point()
p1

p1 + theme_bw(base_size=20, base_family="serif") #you can keep adding to the plot

p2 <- ggplot(data=d, aes(x=fl, fill=fl)) +
  geom_bar()
p2 + coord_flip() + theme_bw(base_size=15, base_family="sans")
```

#### More theme modifications
```{r}
p3 <- ggplot(d, aes(x=displ, y=cty)) +
  geom_point(size=5, shape=21, color="magenta", fill="black") +
  xlab("count") +
  ylab("Fuel") +
  labs(title="Fuel vs. Count", subtitle="subtitle")

p3 + xlim(1,10) + ylim(0, 35)

cols <- viridis(7, option="magma") #specify you want 7 colors for 7 groups
ggplot(d, aes(x=class, y=hwy, fill=class)) +  #fill=class gives color
  geom_boxplot() +
  scale_fill_manual(values=cols)

```

## Multi Panel Plot (`patchwork`)
```{r}
p1+p2+p3
p1/p2/p3
(p1+p2)/p3
```