---
title: "Loading Data"
author: "Alex Takoudes"
date: "2023-01-26"
output: html_document
---

### `read` Functions
* Best format for data is `.csv`
* `header=T` if you have column names
* `sep = ","` seperates data by commans

```{r, eval=F}

#these functions read in datasets
read.table(file="path/to/data.csv, header=TRUE, sep= ",") #header specifies that columns have names
read.csv(file="data.csv", header=T)
```

### `write` Functions
* Create and save a dataset
```{r, eval=F}
write.table(x=varname, file="outputfilename.csv", header=T, sep=",")
```

#### Use RDS Object when only working in R
* `saveRDS()`
* `readRDS()`
* Make sure to put .RDS at the end
* Helpful for large datasets
```{r, eval=F}
saveRDS(my_data, file="filename.RDS")
P <- readRDS("filename.RDS")
```

#### Long vs. Wide Data Formats
* Wide format contains values that **do not** repeat in the ID column
* Long format contains values that **do** repeat in the ID column
```{r, collapse=T}
library(tidyverse)
```

***
#### `pivot_longer()` converts wide to long

```{r}
head(billboard) #wide

b1 <- billboard %>%
  pivot_longer(
    cols = starts_with("wk"), #specify which columns you want to make longer
    names_to = "Week",        #name of new column
    values_to = "Rank",       #name of a new column which will contain values
    values_drop_na = TRUE     #removes any values where the values = NA
    )

glimpse(b1)
```

***
#### `pivot_wider()` convers to wide
* Best for making occupancy matrix

```{r}

glimpse(fish_encounters)

#we want an occupancy matrix where the fish are the rows and the locations are the columns
fish_encounters%>%
  pivot_wider(
    names_from = station, #which column you want to turn into multiple columns
    values_from = seen,   #which column contains the values for the new columns
    values_fill = 0       #all NAs will be 0
    )
```

***
## Loading in Data
* Dyrad: makes research data freely reusable to everybody

```{r}
data <- read.table(
  file="/Users/alextakoudes/BIOL1007A/Data/veysey-babbitt_data_amphibians.csv",
  header=TRUE,
  sep=","
  )

# check how many different species there are!!
table(data$species) #tells us that there's two species

summary(data$mean.hydro)
```

#### `factor()`
* Different type of data