---
title: "Assignment 4"
author: "Alex Takoudes"
date: "2023-01-17"
output: html_document
---

### 1. Matrix with n_dims

Assign to the variable n_dims a single random integer between 3 and 10.
```{r}
set.seed(100)
n_dims <- sample(3:10, 1)
```
Create a vector of consecutive integers from 1 to n_dims2
```{r}
my_vec <- c(1:n_dims^2)
```
Use the sample function to randomly reshuffle these values.
```{r}
random_vec <- sample(my_vec)
```
create a square matrix with these elements.
```{r}
my_mat <- matrix(data=random_vec, nrow=n_dims, ncol=n_dims)
```
print out the matrix.
```{r}
my_mat
```
find a function in r to transpose the matrix
print it out again and note how it has changed
```{r}
my_mat <- t(my_mat)
my_mat
```
calculate the sum and the mean of the elements in the first row and then the last row.
```{r}
#first row
sum(my_mat[1,])
mean(my_mat[1,])

#last row
sum(my_mat[n_dims,])
mean(my_mat[n_dims,])
```
set your code up so you can re-run it to create a matrix of a different size by only changing the n_dims value
```{r}
#done
```

### 2. List Fun

Create a list with the following named elements:

my_matrix, which is a 4 x 4 matrix filled with random uniform values
my_logical which is a 100-element vector of TRUE or FALSE values. You can do this efficiently by setting up a vector of random values and then applying an inequality to it.
my_letters, which is a 26-element vector of all the lower-case letters in random order.
Then, complete the following steps:
```{r}
my_list <- list(my_matrix=matrix(data=runif(16), nrow=4, ncol=4), my_logical= (runif(100) > 0.5), my_letters=sample(letters))
my_list
```

create a new list, which has the element[2,2] from the matrix, the second element of the logical vector, and the second element of the letters vector.
```{r}
new_list <- list(my_list$my_matrix[2,2], my_list$my_logical[[2]], my_list$my_letters[[2]])
new_list
```

use the typeof() function to confirm the underlying data types of each component in this list
```{r}
typeof(new_list[[1]])
typeof(new_list[[2]])
typeof(new_list[[3]])
#or more efficiently...
str(new_list)
```

combine the underlying elements from the new list into a single atomic vector with the c() function.
```{r}
combined_vec <- c(unlist(new_list))
```
what is the data type of this vector?
```{r}
typeof(combined_vec)
```

Create an empty data frame with two variables (= columns) and 26 observations (= rows) below:

call the first column/variable my_unifs and fill it with 26 random uniform values from 0 to 10

call the second variable my_letters and fill it with 26 capital letters in random order.

for the first variable, use a single line of code in R to select 4 random rows and replace the numerical values in those rows with NA.

for the first variable, write a single line of R code to identify which rows have the missing values.
re-order the entire data frame to arrange the second variable in alphabetical order.

calculate the column mean for the first variable.
rename the second column ‘my_LETTERS’