---
title: "Assignment 9"
author: "Alex Takoudes"
date: "2023-01-30"
output: html_document
---

##### 1. Using a for loop and an if statement, write a function to calculate the number of zeroes in a numeric vector. Before entering the loop, set up a vector of integers with some zero values. Then, set up a counter variable counter <- 0. Inside the loop, add 1 to counter each time you have a zero in the vector. Finally, use return(counter) for the output.

```{r}
my_vec <- sample(c(0:5), size=100, replace=T) #initialize a vector to be sampled
my_vec

zero_counter <- function(vector=my_vec){
  counter <- 0                        #initialize the counter
  for(i in my_vec){                   #get a little fancy and just iterate through the actual values of my_vec
    if(i==0){counter <- counter + 1}  #if that value is 0, add 1 to the counter
  }
  return(counter)
}

zero_counter()
```

##### 2. Write a function that takes as input two integers representing the number of rows and columns in a matrix. The output is a matrix of these dimensions in which each element is the product of the row number x the column number.

```{r}
multiply <- function(rows, cols){
  matrix <- matrix(nrow=rows, ncol=cols)
  for (i in 1:rows) {
    for (j in 1:cols) {
      matrix[i,j] <- i*j
    } # end of column j loop
  } # end or row i loop
  return(matrix)
}

multiply(2, 2)
multiply(5, 3)
```

##### 3. Use a for loop to repeat your previous functions from Weekly Assignment 2 (which simulate a data frame and extract a p-value) 100 times. Store the results in a data frame that has 1 column indicating the replicate number and 1 column for the p-values (a total of 2 columns and 100 rows).

```{r}

##################################################
# FUNCTION: datagen 
# generates a two column data frame of data.
# inputs:
#   c_mean: mean value for normal distribution of control data
#   c_sd: standard deviation value for normal distribution of control data
#   t_mean: mean value for normal distribution of treatment data
#   t_sd: standard deviation value for normal distribution of treatment data
#   nrows: number of rows of data
# outputs: 2 column matrix nrows long with half control data and half treatment data
#------------------------------------------------- 

datagen <- function(c_mean=10, c_sd=1.5, t_mean=45, t_sd=2, nrows=50){
  dFrame <- data.frame(
    Type = rep(c("Control", "Treatment"),each=nrows/2), 
    Value =  c(control = rnorm(nrows/2, mean=c_mean, sd=c_sd),
               treatment = rnorm(nrows/2, mean=t_mean, sd=t_sd)
               )
    )
  return(dFrame)
}
##################################################

##################################################
# FUNCTION: psearch
# returns the pvalue of a given two column data frame
# inputs:
#   df: data frame to be analyzed
# outputs: p value of two column data frame
#------------------------------------------------- 
psearch <- function(df=NULL){
  if(is.null(df)){
    df <- datagen()
  }
  test <- aov(Value ~ Type, data=df)
  testSummary <- summary(test)
  pvalue <- testSummary[[1]][["Pr(>F)"]][1]
  return(pvalue)
}
##################################################

##################################################
# FUNCTION: dataloop
# generates a data frame containing 100 simulated values for P from datagen() and psearch()
# inputs:
#   x: number of iterations
# outputs: 2x100 dataframe with replicate numbers and p values
#------------------------------------------------- 
dataloop <- function(x=100){
  pvals <- vector("double", x)    #initialize pvals as a 100 item long vector
  counts <- vector("integer", x)  #initialize counts as a 100 item long vector
  for(i in 1:100){
    pvals[i] <- psearch()
    counts[i] <- i
  }
  frame <- data.frame(counts, pvals) #turn the vectors into a data frame
  return(frame)
}
##################################################

answer <- dataloop()
tail(answer)
```