---
title: "Assignment 4"
author: "Alex Takoudes"
date: "2023-01-17"
output: html_document
---

### 1. Matrix with n_dims

Assign to the variable n_dims a single random integer between 3 and 10.
```{r}
set.seed(100)
n_dims <- sample(3:10, 1)
```
Create a vector of consecutive integers from 1 to n_dims2
```{r}
my_vec <- c(1:n_dims^2)
```
Use the sample function to randomly reshuffle these values.
```{r}
random_vec <- sample(my_vec)
```
create a square matrix with these elements.
```{r}
my_mat <- matrix(data=random_vec, nrow=n_dims, ncol=n_dims)
```
print out the matrix.
```{r}
my_mat
```
find a function in r to transpose the matrix
print it out again and note how it has changed
```{r}
my_mat <- t(my_mat)
my_mat
```
calculate the sum and the mean of the elements in the first row and then the last row.
```{r}
#first row
sum(my_mat[1,])
mean(my_mat[1,])

#last row
sum(my_mat[n_dims,])
mean(my_mat[n_dims,])
```
set your code up so you can re-run it to create a matrix of a different size by only changing the n_dims value
```{r}
#done
```

### 2. 

Create a list with the following named elements:

my_matrix, which is a 4 x 4 matrix filled with random uniform values
my_logical which is a 100-element vector of TRUE or FALSE values. You can do this efficiently by setting up a vector of random values and then applying an inequality to it.
my_letters, which is a 26-element vector of all the lower-case letters in random order.
Then, complete the following steps:
```{r}
my_list <- list(my_matrix=matrix(data=runif(16), nrow=4, ncol=4), my_logical=(a>0.5(rep(runif(10)), 
```



create a new list, which has the element[2,2] from the matrix, the second element of the logical vector, and the second element of the letters vector.
use the typeof() function to confirm the underlying data types of each component in this list
combine the underlying elements from the new list into a single atomic vector with the c() function.
what is the data type of this vector?