---
title: "Assignment 4"
author: "Alex Takoudes"
date: "2023-01-17"
output: html_document
---

### 1. Matrix with n_dims

##### a. Assign to the variable n_dims a single random integer between 3 and 10.
```{r}
set.seed(1)
n_dims <- sample(3:10, size=1)
```
##### b. Create a vector of consecutive integers from 1 to n_dims2
```{r}
my_vec <- c(1:n_dims^2)
```
##### c. Use the sample function to randomly reshuffle these values.
```{r}
random_vec <- sample(my_vec)
```
##### d. Create a square matrix with these elements.
```{r}
my_mat <- matrix(data=random_vec, nrow=n_dims, ncol=n_dims)
```
##### e. Print out the matrix.
```{r}
my_mat
```
##### f. Find a function in r to transpose the matrix. Print it out again and note how it has changed
```{r}
my_mat <- t(my_mat) 
my_mat #rows become columns
```
##### h. Calculate the sum and the mean of the elements in the first row and then the last row.
```{r}
#first row
sum(my_mat[1,]); mean(my_mat[1,])

#last row
sum(my_mat[n_dims,]); mean(my_mat[n_dims,])
```
##### i. set your code up so you can re-run it to create a matrix of a different size by only changing the n_dims value
```{r}
#done
```

### 2. Messing With Lists

##### a. Create a list with the following named elements:
* my_matrix, which is a 4 x 4 matrix filled with random uniform values
* my_logical which is a 100-element vector of TRUE or FALSE values. You can do this efficiently by setting up a vector of random values and then applying an inequality to it.
* my_letters, which is a 26-element vector of all the lower-case letters in random order.
```{r}
my_list <- list(
  my_matrix=matrix(data=runif(16), nrow=4, ncol=4), 
  my_logical= (runif(100) > 0.5), 
  my_letters=sample(letters)
  )
my_list
```
##### b. Create a new list, which has the element[2,2] from the matrix, the second element of the logical vector, and the second element of the letters vector.
```{r}
new_list <- list(
  my_list$my_matrix[2,2],
  my_list$my_logical[[2]], 
  my_list$my_letters[[2]]
  )
new_list
```

##### c. Use the typeof() function to confirm the underlying data types of each component in this list
```{r}
typeof(new_list[[1]])
typeof(new_list[[2]])
typeof(new_list[[3]])
#or more efficiently...
str(new_list)
```

##### d. Combine the underlying elements from the new list into a single atomic vector with the c() function.
```{r}
combined_vec <- unlist(new_list)
```
##### e. What is the data type of this vector?
```{r}
typeof(combined_vec)
```

### 3. Data Frames

##### a. Create an empty data frame with two variables (= columns) and 26 observations (= rows) below:
```{r}
dFrame <- data.frame(matrix(ncol=2, nrow=26))
tail(dFrame)
```

##### b. Call the first column/variable my_unifs and fill it with 26 random uniform values from 0 to 10
```{r}
my_unifs <- runif(26, min=0, max=10)
dFrame <- data.frame(my_unifs, NA)
tail(dFrame)
```

##### c. Call the second variable my_letters and fill it with 26 capital letters in random order.
```{r}
my_letters <- sample(LETTERS)
dFrame <- data.frame(my_unifs, my_letters)
tail(dFrame)
```

##### d. For the first variable, use a single line of code in R to select 4 random rows and replace the numerical values in those rows with NA.
```{r}
dFrame$my_unifs[sample(1:nrow(dFrame), 4)] <- NA
dFrame
```

##### e. For the first variable, write a single line of R code to identify which rows have the missing values.

```{r}
which(is.na(dFrame$my_unifs))
```

##### f. Re-order the entire data frame to arrange the second variable in alphabetical order.
```{r}
dFrame <- dFrame[order(dFrame$my_letters),]
dFrame
```

##### g. Calculate the column mean for the first variable. 
```{r}
mean(dFrame$my_unifs, na.rm=TRUE)
```

##### h. Rename the second column ‘my_LETTERS'
```{r}
colnames(dFrame)[2] <- "my_LETTERS"
tail(dFrame)
```

